---
title: 'Police Data Challenge'
author: "Matt Sterkel"
output: html_document
---
read in bike theft data

```{r, warning=FALSE,message=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)

seat<-read.csv("https://raw.githubusercontent.com/ClassicSours/policedatachallenge/master/biketheft.csv")

```

subset data to year 2015,and selected variables

```{r}
by2015<-filter(seat,YYYY == 2015) %>% select(Event.Clearance.Description,CAD.CDW.ID,YYYY,MM,DD,
                                             TIME,Census.Tract,Latitude,Longitude) %>%
              mutate(month=str_pad(MM,2,pad="0"),
                     day=str_pad(DD,2,pad="0"),  
                     date = ymd(paste0(YYYY,month,day)))

```

read in event data

```{r, warning=FALSE, message=FALSE}
library(readxl)
library(tidyr)

eventData<-read_xlsx("../police_Data/Festivals and Sports Events1.xlsx",col_names = TRUE)%>% separate(Date,c("YYYY","MM","DD"),"-")%>%
              mutate(month=str_pad(MM,2,pad="0"),
                     day=str_pad(DD,2,pad="0"),  
                     date = ymd(paste0(YYYY,month,day)))
  
```



determine if the theft occured on the same day as an event (anywhere)
```{r}
#convert dates to #s 
#add as column to by2015 data
 by2015<-by2015%>%mutate(matchDates=by2015$date %in% eventData$date) 
 
 
```


determine if theft occured in 1 mile radius of event location using lat/long
used conversion of 1 degree of long/ 69.2 miles

```{r}
latLow<-c(eventData$Lattitude-0.01445)
latHigh<-c(eventData$Lattitude+0.01445)
longLow<-c(eventData$Longitude-0.01445)
longHigh<-c(eventData$Longitude+0.01445)

 
m<-rep(0,nrow(by2015))
     
for(i in 1:nrow(by2015)){
  if(by2015$matchDate[i] == TRUE){
    if(any(by2015$Latitude[i]>latLow   & by2015$Latitude[i]<latHigh & 
           by2015$Longitude[i]>longLow & by2015$Longitude[i]<longHigh))
     m[i]<-1
    }
}

mean(m) 

by2015<-by2015%>%mutate(m)

```

graph theft and events on y-axis and time over x-axis

```{r, message=FALSE,warning=FALSE}
  library(reshape2)
 ggplot(NULL, aes(month))+ 
  geom_bar(aes(fill = "events"), data = eventData, alpha = 0.5)+
  geom_bar(aes(fill = "bike thefts"), data = by2015,alpha=0.3)+
   scale_fill_manual(name="occurance type",
                   values = c('bike thefts' = '#f67e55','events' = 'black')) 

sumdata<-by2015%>%group_by(date)%>%summarize(n=n())
sumdataEvent<-eventData%>%group_by(date)%>%summarize(n=n())



ggplot()+
  geom_point(data=sumdata, aes(x=date,y=n),color="red")+
  ylab("number of occurances")+
    geom_smooth(data=sumdata, aes(x=date,y=n,color="Bike thefts"),se=FALSE)+
  geom_point(data=sumdataEvent,aes(x=date,y=n),color="blue")+
    geom_smooth(data=sumdataEvent, aes(x=date,y=n,color="events"),se=FALSE)+
scale_color_manual(name="occurance type",
                   values = c('Bike thefts' = 'red','events' = 'blue')) 



```

Proportion of bike thefts during events by month(1-12)


```{r}
sumdata<-by2015%>%group_by(month)%>%summarize(n=n(),xm=sum(m),p=paste0(round(xm/n,2)*100,"%"))
sumdata$p
```

we see that during the summer months, the proportion of bike thefts during events increases.

```{r}

ggplot(by2015, aes(x=month,fill=as.factor(m)))+
  geom_bar()+
  scale_fill_manual(name="",
                    values = c("#aaa4a1","#f67e55"),
                    labels=c("Bike thefts","Bike thefts during event"))
```


use pie chart showing proportion of bike thefts and those thefts that are durring event(time) And within radius of event



```{r}
chart <- table(by2015$m)

  names(chart)<-c("no event","event")
    pie(chart, labels = paste0(names(chart), ' (', round(prop.table(chart)*100, 2), "%)"),col=c("#aaa4a1","#f67e55"),main="Proportion of bike thefts during events")
```

